Adversarial Training of Neural Networks\\ against Systematic Uncertainty

### 2016-08-02 paper.tex
save

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
nuisance parameter $\lambda \in \mathbb{R}$ whose values define a continuously
parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
$Y$ induce together a joint probability distribution $p(X,Y|\lambda)$, where the
conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
finite set $\{ x_i, y_i, \lambda_i \}_{i=1}^N$ of realizations
$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function
$f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of parameters $\theta_f$
(e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
classes) minimizing  a loss ${\cal L}_f(\theta_f)$, with the constraint that $f(X_\lambda ;
\theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
which remains unknown at test time. More specifically, we aim at building $f$
such that in the ideal case
\begin{equation}\label{eqn:criterion-true}
f(X_{\lambda_i}(\omega) ; \theta_f) = f(X_{\lambda_j}(\omega) ; \theta_f)
\end{equation} for any
sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
nuisance parameter.

@@

In this work, we repurpose adversarial training as a means to regularize the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
particular, we pit $f$ against an adversary regression model $r(\cdot ;
\theta_r) : \mathbb{R} \mapsto \mathbb{R}$ of parameters $\theta_r$ and
associated loss ${\cal L}_r(\theta_r)$, taking as input realizations of $f(X_\lambda; \theta_f)$
and producing as output a prediction $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
of the nuisance parameter. If $f(X_\lambda; \theta_f)$ varies with $\lambda$,
then the corresponding correlation can be captured by $r$. By contrast, if
$f(X_\lambda; \theta_f)$ is invariant with $\lambda$ as we require, then $r$
should perform poorly. Training $f$ such that it minimizes the performance of $r$
should therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.

@@

\glnote{What are the necessary conditions on ${\cal L}_r$ to imply
Eqn.~\ref{eqn:criterion} at the saddle point? We should clarify and prove that
formally! There may also be assumptions required on the learning problem itself.}

>>>>

Let assume a probability space $(\Omega, {\cal F}, P)$, where $\Omega$ is a
sample space, ${\cal F}$ is a set of events and $P$ is a probability measure.
Let consider the multivariate random variables $X_\lambda: \Omega \mapsto
\mathbb{R}^p$ and $Y: \Omega \mapsto {\cal Y}$, where $X_\lambda$ depends on a
nuisance parameter $\lambda \in \mathbb{R}$ whose values define a continuously
parameterized family of its systematic uncertainties. That is, $X_\lambda$ and
$Y$ induce together a joint probability distribution $p(X,Y|\lambda)$, where the
conditional on $\lambda$ denotes $X_\lambda$. For training, let further assume a
finite set $\{ x_i, y_i, \lambda_i \}_{i=1}^N$ of realizations
$X_{\lambda_i}(\omega_i), Y(\omega_i)$, for $\omega_i \in \Omega$ and known
values $\lambda_i$ of the nuisance parameter. Our goal is to learn a function
$f(\cdot;\theta_f) : \mathbb{R}^p \mapsto {\cal Y}$ of parameters $\theta_f$
(e.g., a neural network-based classifier if ${\cal Y}$ is a finite set of
classes) and minimizing  a loss ${\cal L}_f(\theta_f)$. In addition, we require
the constraint that $f(X_\lambda ;
\theta_f)$ should be robust to the value of the nuisance parameter $\lambda$ --
which remains unknown at test time. More specifically, we aim at building $f$
such that in the ideal case
\begin{equation}\label{eqn:criterion-true}
f(X_{\lambda_i}(\omega) ; \theta_f) = f(X_{\lambda_j}(\omega) ; \theta_f)
\end{equation} for any
sample $\omega \in \Omega$ and any $\lambda_i, \lambda_j$ pair of values of the
nuisance parameter.

@@

In this work, we repurpose adversarial training as a means to regularize the
predictive model $f$ in order to satisfy Eqn.~\ref{eqn:criterion}. In
particular, we pit $f$ against an adversary regression model $r(\cdot ;
\theta_r) : \mathbb{R} \mapsto \mathbb{R}$ of parameters $\theta_r$ and
associated loss ${\cal L}_r(\theta_r)$, taking as input realizations of $f(X_\lambda; \theta_f)$
and producing as output predictions $\smash{\hat\lambda} = r(f(X_\lambda); \theta_f)$
of the nuisance parameter. If $f(X_\lambda(\omega); \theta_f)$ varies with $\lambda$,
then the corresponding correlation can be captured by $r$. By contrast, if
$f(X_\lambda(\omega); \theta_f)$ is invariant with $\lambda$ as we require, then $r$
should perform poorly. Training $f$ such that it minimizes the performance of $r$
should therefore acts as a regularization towards Eqn.~\ref{eqn:criterion}.

@@

\glnote{Explain this can be done simply using SGD.}
\glnote{What are the necessary conditions on ${\cal L}_r$ to imply
Eqn.~\ref{eqn:criterion} at the saddle point? We should clarify and prove that
formally! There may also be assumptions required on the learning problem itself.}
